---
title: "OAK vs. TOR"
output: html_document
date: "2023-03-02"
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages:
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(baseballr)
library(plotly)
library(CalledStrike)
library(knitr)
library(kableExtra)
library(rpart)
```

Load data:
```{r, include=FALSE}
library(readr)
OAK_TOR <- read_csv("C:/Users/jacki/Desktop/SABR/2023/Group A/Group A/Nelson 4-16-22/OAK at TOR 4-16-22.csv")
View(OAK_TOR)

# make variable for radius of baseball IN FEET
b_rad <- 2.9/2/12
```

```{r, message=FALSE, warning=FALSE}
# filter out columns with majority NAs, or variables not needed for this analysis
clean <- OAK_TOR %>% 
  dplyr::select(pitch_type:pitcher...8, description, zone:type, balls:plate_z, outs_when_up:inning_topbot, sz_top, sz_bot, effective_speed, release_spin_rate, 
                game_pk, release_pos_y, at_bat_number:post_bat_score, spin_axis:delta_run_exp)

# add an id for each observation
clean <- clean %>% 
  mutate(id = 1:277)

# filter to pitches of interest and add in sz_l and sz_r
OAK_TOR_filtered <- clean %>% 
  filter(description %in% c("ball", "blocked_ball", "called_strike")) %>% 
  mutate(sz_l = -8.5/12, sz_r = 8.5/12)
```

### Finding the incorrect calls:
```{r, fig.align='center'}
long <- OAK_TOR_filtered %>% 
  mutate(new_sz_top = sz_top + b_rad,
         new_sz_bot = sz_bot - b_rad,
         new_sz_l = sz_l - b_rad,
         new_sz_r = sz_r + b_rad)
long <- long %>% 
  dplyr::select(-c(game_year:pfx_z))

calc_long <- long %>% 
  mutate(true = case_when(plate_z > new_sz_top | plate_z < new_sz_bot ~ "ball",
                          plate_x > new_sz_r | plate_x < new_sz_l ~ "ball",
                          TRUE ~ "strike"))
table(calc_long$true)

graph <- calc_long %>% 
  ggplot(aes(x = plate_x, y = plate_z, color = true)) +
  geom_point() +
  labs(title = "Call result from the code") +
  theme_bw() +
  add_zone() + coord_fixed(ratio = 1)
graph
#ggplotly(graph)

calc_long %>% 
  ggplot(aes(x = plate_x, y = plate_z, color = type)) +
  geom_point() +
  labs(title = "Umpire's call") +
  theme_bw() +
  add_zone() + coord_fixed(ratio = 1)
```

```{r}
identify <- calc_long %>% 
  mutate(call = case_when((type == "B" & true == "ball") ~ "pass",
                          (type == "S" & true == "strike") ~ "pass",
                          TRUE ~ "fail"))
table(identify$call)
```

All pitches that were called_strike, ball, and blocked_ball
```{r}
fill <- which(identify$call == "fail")

identify %>% 
  dplyr::select(-des) %>% # took out this variable cause it was long
  kable() %>% 
  kable_styling("striped") %>% 
  row_spec(fill, background = "skyblue") %>% 
  scroll_box(height = "500px", width = "100%")
```

Only the ones identified as incorrect:
```{r}
incorrect <- identify %>% 
  filter(call == "fail")
incorrect %>% 
  dplyr::select(-des) %>% # took out this variable cause it was long
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "500px", width = "100%")
```

Finding counts of incorrect calls:
```{r}
incorrect %>% 
  group_by(balls, strikes) %>% 
  summarize(n())
```
Most of the incorrect calls were made during a 0-0 count

## Predictive Modeling

### Predicting Incorrect Calls
```{r}
model <- rpart(type ~ outs_when_up + inning + home_score + away_score + balls + strikes, data = identify)
predictions <- as.data.frame(predict(model))
predictions <- predictions %>% 
  mutate(predict_call = case_when(B > S ~ "ball",
                                  S > B ~ "strike"))
table(predictions$predict_call)
```
This is really close to what the true amount of balls and strikes is

Compare predicted call to actual call
```{r}
# add predictions to the data frame
pred_m <- calc_long %>% 
  mutate(predict_call = predictions$predict_call)

# find the incorrect calls
identify_pm <- pred_m %>% 
  mutate(call = case_when((type == "B" & true == "ball") ~ "pass",
                          (type == "S" & true == "strike") ~ "pass",
                          TRUE ~ "fail"))
table(identify_pm$call)
```
Still produces the same number of fail/pass

### Predicting Individual Play Results

**First draft:**
```{r}
clean_added <- clean %>% 
  mutate(ump_call = case_when(id %in% incorrect$id ~ "incorrect",
                              TRUE ~ "correct"))

model_outcome <- rpart(description ~ ump_call + outs_when_up + inning + bat_score + fld_score + balls + strikes, data = clean_added)
predictions_outcome <- as.data.frame(predict(model_outcome))

# get the column names of the largest values
predictions_outcome$highest<- (colnames(predictions_outcome)[max.col(predictions_outcome)])

# add in to data frame
clean_added <- clean_added %>% 
  mutate(pred_outcome = predictions_outcome$highest)
```

```{r}
fill2 <- which(clean_added$ump_call == "incorrect")
clean_added %>% 
  dplyr::select(-des) %>% 
  kable %>% 
  kable_styling("striped") %>% 
  row_spec(fill2, background = "skyblue") %>% 
  scroll_box(height = "500px")
```


**Updated version:**
```{r, include=FALSE}
correct <- OAK_TOR %>% 
  mutate(id = 1:277)

correct <- correct %>% 
  filter(!(id %in% identify$id))

correct <- correct %>% 
  dplyr::select(pitch_type:pitcher...8, description, zone:type, balls:plate_z, outs_when_up:inning_topbot, sz_top, 
                sz_bot, effective_speed, release_spin_rate, 
                game_pk, release_pos_y, at_bat_number:post_bat_score, spin_axis:id)

correct <- correct %>% 
  mutate(sz_l = NA, sz_r = NA, new_sz_top = NA, new_sz_bot = NA, new_sz_l = NA, new_sz_r = NA,
         true = description, call = "pass")

correct <- correct %>% 
  dplyr::select(-game_year, -pfx_x, -pfx_z)

all_added <- rbind(identify, correct)
```

```{r}
model_outcome2 <- rpart(description ~ call + true + outs_when_up + inning + bat_score + fld_score + balls + 
                          strikes + plate_x + plate_z, 
                       data = all_added)
predictions_outcome2 <- as.data.frame(predict(model_outcome2))

# get the column names of the largest values
predictions_outcome2$highest<- (colnames(predictions_outcome2)[max.col(predictions_outcome2)])

# add in to data frame
all_added <- all_added %>% 
  mutate(pred_outcome = predictions_outcome2$highest)
```


Full data with incorrect pitches highlighted:
```{r}
fill3 <- which(all_added$ump_call == "incorrect")
all_added %>% 
  dplyr::select(-des) %>% 
  kable %>% 
  kable_styling("striped") %>% 
  row_spec(fill3, background = "skyblue") %>% 
  scroll_box(height = "500px")
```


Just the incorrect pitches with how umpire called it and the predicted outcome:
```{r}
all_added %>% 
  filter(call == "fail") %>% 
  dplyr::select(id, description, pred_outcome) %>% 
  kable %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "500px")
```



### Visualization of Model Predictions {.tabset}

#### All 
```{r, fig.align='center'}
all_added %>% 
  ggplot(aes(x = plate_x, y = plate_z, color = pred_outcome)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  add_zone() + coord_fixed(ratio = 1)
```

#### Faceted
```{r, fig.align='center'}
all_added %>% 
  ggplot(aes(x = plate_x, y = plate_z, color = pred_outcome)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  add_zone() + coord_fixed(ratio = 1) +
  facet_wrap(~pred_outcome)
```

### {-}

**Splitting this game into train and test data:**

```{r, include=FALSE}
train <- clean %>% 
  filter(!(id %in% incorrect$id)) %>% 
  mutate(ump_call = "correct")

train_calc <- train %>% 
  filter(id %in% calc_long$id)
train_calc_b <- calc_long %>% 
  filter(id %in% train$id)
train_calc <- train_calc %>% 
  mutate(true = train_calc_b$true)

train_other <- train %>% 
  filter(!(id %in% train_calc$id)) %>% 
  mutate(true = description)
train <- rbind(train_calc, train_other)

test <- clean %>% 
  filter(id %in% incorrect$id) %>% 
  mutate(ump_call = "incorrect")
```

```{r}
model_outcome3 <- rpart(description ~ outs_when_up + inning + bat_score + fld_score + balls + #true + ump_call +
                          strikes + plate_x + plate_z, 
                       data = train)
predictions_outcome3 <- as.data.frame(predict(model_outcome3, newdata = test))

# get the column names of the largest values
predictions_outcome3$highest<- (colnames(predictions_outcome3)[max.col(predictions_outcome3)])

# add in to data frame
test$pred_outcome <- predictions_outcome3$highest
```

```{r}
test %>% 
  dplyr::select(id, description, balls, strikes, outs_when_up, inning, bat_score, fld_score, pred_outcome) %>% 
  kable %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "500px")
```

```{r, fig.align='center'}
graph_3 <- test %>% 
  ggplot(aes(x = plate_x, y = plate_z, color = pred_outcome)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  add_zone() + coord_fixed(ratio = 1)
ggplotly(graph_3)
```

This is not a perfect model - it is not possible for an umpire to call a pitch (not hit) a foul, and some of the strikes are outside of the strike zone but they are close (since this is a generalized strike zone, not specific to each pitch, it is possible that the true strike zone is slightly different)





