---
title: "TEX vs. MIA"
output: html_document
date: "2023-03-04"
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages:
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(baseballr)
library(plotly)
library(CalledStrike)
library(knitr)
library(kableExtra)
library(rpart)
```

Load data:
```{r, include=FALSE}
library(readr)
TEX_MIA <- read_csv("C:/Users/jacki/Desktop/SABR/2023/Group B/Group B/Libka 9-12-22/TEX at MIA 9-12-22.csv")
View(TEX_MIA)

# make variable for radius of baseball IN FEET
b_rad <- 2.9/2/12
```

```{r, message=FALSE, warning=FALSE}
# filter out columns with majority NAs, or variables not needed for this analysis
clean_tm <- TEX_MIA %>% 
  dplyr::select(pitch_type:pitcher...8, description, zone:type, balls, strikes, plate_x, plate_z, outs_when_up:inning_topbot, sz_top, sz_bot, effective_speed, 
                release_spin_rate, game_pk, release_pos_y, at_bat_number:post_bat_score, spin_axis:delta_run_exp) %>% 
  mutate(id = 1:273)

# filter to pitches of interest and add in sz_l and sz_r
TEX_MIA_filtered <- clean_tm %>% 
  filter(description %in% c("ball", "blocked_ball", "called_strike")) %>% 
  mutate(sz_l = -8.5/12, sz_r = 8.5/12)
```

### Finding the incorrect calls:
```{r}
long_tm <- TEX_MIA_filtered %>% 
  mutate(new_sz_top = sz_top + b_rad,
         new_sz_bot = sz_bot - b_rad,
         new_sz_l = sz_l - b_rad,
         new_sz_r = sz_r + b_rad)

calc_long_tm <- long_tm %>% 
  mutate(true = case_when(plate_z > new_sz_top | plate_z < new_sz_bot ~ "ball",
                          plate_x > new_sz_r | plate_x < new_sz_l ~ "ball",
                          TRUE ~ "strike"))
table(calc_long_tm$true)
```
The predicted number of balls and strikes is very close to the actual number.

```{r, fig.align='center'}
calc_long_tm %>% 
  ggplot(aes(x = plate_x, y = plate_z, color = true)) +
  geom_point() +
  labs(title = "Call result from the code") +
  theme_bw() +
  add_zone() + coord_fixed(ratio = 1)

calc_long_tm %>% 
  ggplot(aes(x = plate_x, y = plate_z, color = type)) +
  geom_point() +
  labs(title = "Umpire's call") +
  theme_bw() +
  add_zone() + coord_fixed(ratio = 1)
```


```{r}
identify_tm <- calc_long_tm %>% 
  mutate(call = case_when((type == "B" & true == "ball") ~ "pass",
                          (type == "S" & true == "strike") ~ "pass",
                          TRUE ~ "fail"))
table(identify_tm$call)
```

All pitches that were called_strike, ball, and blocked_ball
```{r}
fill4 <- which(identify_tm$call == "fail")

identify_tm %>% 
  dplyr::select(-des) %>% # took out this variable cause it was long
  kable() %>% 
  kable_styling("striped") %>% 
  row_spec(fill4, background = "skyblue") %>% 
  scroll_box(height = "500px", width = "100%")
```

Only the ones identified as incorrect:
```{r}
incorrect_tm <- identify_tm %>% 
  filter(call == "fail")
incorrect_tm %>% 
  dplyr::select(-des) %>% # took out this variable cause it was long
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(height = "500px", width = "100%")
```

Finding counts of incorrect calls:
```{r}
incorrect_tm %>% 
  group_by(balls, strikes) %>% 
  summarize(n())
```
Most of the incorrect calls were made during a 0-0 count, but the spread is pretty even.

## Predictive Modeling

### Predicting Incorrect Calls

```{r}
model_tm <- rpart(type ~ outs_when_up + inning + home_score + away_score + balls + strikes, data = identify_tm)
predictions_tm <- as.data.frame(predict(model_tm))
predictions_tm <- predictions_tm %>% 
  mutate(predict_call = case_when(B > S ~ "ball",
                                  S > B ~ "strike"))
table(predictions_tm$predict_call)
```
Less accurate than previous way

Compare predicted call to actual call
```{r}
# add predictions to the data frame
pred_m_tm <- calc_long_tm %>% 
  mutate(predict_call = predictions_tm$predict_call)

# find the incorrect calls
identify_pm_tm <- pred_m_tm %>% 
  mutate(call = case_when((type == "B" & true == "ball") ~ "pass",
                          (type == "S" & true == "strike") ~ "pass",
                          TRUE ~ "fail"))
table(identify_pm_tm$call)
```

### Predicting Individual Play Results

```{r, include=FALSE}
correct_tm <- TEX_MIA %>% 
  dplyr::select(pitch_type:pitcher...8, description, zone:type, balls, strikes, plate_x, plate_z, outs_when_up:inning_topbot, sz_top, sz_bot, effective_speed, 
                release_spin_rate, game_pk, release_pos_y, at_bat_number:post_bat_score, spin_axis:delta_run_exp) %>% 
  mutate(id = 1:273)

correct_tm <- correct_tm %>% 
  filter(!(id %in% identify_tm$id)) %>% 
  mutate(sz_l = NA, sz_r = NA, new_sz_top = NA, new_sz_bot = NA, new_sz_l = NA, new_sz_r = NA,
         true = description, call = "pass")

all_added_tm <- rbind(identify_tm, correct_tm)
```

```{r}
model_outcome_tm <- rpart(description ~ call + true + outs_when_up + inning + bat_score + fld_score + balls + 
                          strikes + plate_x + plate_z, data = all_added_tm)
predictions_outcome_tm <- as.data.frame(predict(model_outcome_tm))

# get the column names of the largest values
predictions_outcome_tm$highest<- (colnames(predictions_outcome_tm)[max.col(predictions_outcome_tm)])

# add in to data frame
all_added_tm <- all_added_tm %>% 
  mutate(pred_outcome = predictions_outcome_tm$highest)
```


Full data with incorrect pitches highlighted:
```{r}
fill5 <- which(all_added_tm$call == "fail")
all_added_tm %>% 
  dplyr::select(-des) %>% 
  kable %>% 
  kable_styling("striped") %>% 
  row_spec(fill5, background = "skyblue") %>% 
  scroll_box(height = "500px")
```

Just the incorrect pitches and the predicted outcome:

```{r}
all_added_tm %>% 
  filter(call == "fail") %>% 
  dplyr::select(id, description, pred_outcome) %>% 
  kable %>% 
  kable_styling("striped") 
```

**Splitting this game into train and test data:**

This method does NOT work well for this data frame.

```{r, include=FALSE}
train_tm <- clean_tm %>% 
  filter(!(id %in% incorrect_tm$id)) %>% 
  mutate(ump_call = "correct")

train_calc_tm <- train_tm %>% 
  filter(id %in% calc_long_tm$id)
train_calc_c <- calc_long_tm %>% 
  filter(id %in% train_tm$id)
train_calc_tm <- train_calc_tm %>% 
  mutate(true = train_calc_c$true)

train_other_tm <- train_tm %>% 
  filter(!(id %in% train_calc_tm$id)) %>% 
  mutate(true = description)
train_tm <- rbind(train_calc_tm, train_other_tm)

test_tm <- clean_tm %>% 
  filter(id %in% incorrect_tm$id) %>% 
  mutate(ump_call = "incorrect")
```

```{r, include=FALSE}
model_outcome_tm2 <- rpart(description ~ outs_when_up + inning + bat_score + fld_score + balls + #true + ump_call +
                          strikes + plate_x + plate_z, data = train_tm)
predictions_outcome_tm2 <- as.data.frame(predict(model_outcome_tm2, newdata = test_tm))

# get the column names of the largest values
predictions_outcome_tm2$highest<- (colnames(predictions_outcome_tm2)[max.col(predictions_outcome_tm2)])

# add in to data frame
test_tm$pred_outcome <- predictions_outcome_tm2$highest
```

