---
title: "Visualizations"
output: html_document
date: "2023-03-07"
---

<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Open+Sans:wght@300&family=Source+Sans+Pro&display=swap" rel="stylesheet">
<style>

body{
font-family: 'Source Sans Pro', sans-serif;
}

</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(CalledStrike)
library(knitr)
library(kableExtra)
library(baseballr)
```

These are not all of the graphs, there are some in the original files for each of the games. Here are the links to the [OAK v. TOR](https://htmlpreview.github.io/?https://github.com/jajovanovic/SABR_Case/blob/main/OAK_v_TOR.html) game and the [TEX v. MIA](https://htmlpreview.github.io/?https://github.com/jajovanovic/SABR_Case/blob/main/TEX_v_MIA.html) game. 

## OAK v. TOR

### Win / Loss
```{r, fig.align='center'}
win_loss_ot <- data.frame(team = c("OAK", "OAK", "TOR", "TOR"),
                       record = c(60, 102, 92, 70),
                       result = c("win", "loss", "win", "loss"),
                       season = c(2022, 2022, 2022, 2022))

win_loss_ot %>% 
  group_by(team) %>% 
  ggplot(aes(x = factor(team), y = record, fill = result)) +
  geom_col(position = "dodge") + 
  geom_text(aes(label = record), vjust = -.5, position = position_dodge(.9), family = "serif") +
  labs(x = "Team", y = "Record", fill = "Game Result") +
  #ylim(0, 110) +
  theme_bw() +
  scale_fill_manual(values = c("navyblue", "cornflowerblue")) +
  theme(text=element_text(family = "serif")) 
```



### Payroll

```{r, message=FALSE, warning=FALSE, fig.align='center'}
payroll_ot <- data.frame(team = c("OAK", "TOR"),
                      season = c(2022, 2022),
                      payroll = c(48443900, 177071670))

payroll_ot %>% 
  ggplot(aes(x = factor(team), y = payroll, fill = team)) +
  geom_col(position = "dodge") +
  labs(title = "MLB Payrolls", subtitle = "Oakland Athletics & Toronto Blue Jays", 
       x = "Team", y = "Payroll") +
  theme_bw() +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("springgreen4", "darkblue")) +
  theme(text=element_text(family = "serif")) +
  theme(legend.position = "false")
```


## TEX v. MIA

### Win / Loss

```{r, fig.align='center'}
win_loss <- data.frame(team = c("TEX", "TEX", "MIA", "MIA", "TEX", "TEX", "MIA","MIA", "TEX", "TEX", "MIA", "MIA"),
                       record = c(68, 94, 69, 93, 60, 102, 67, 95, 78, 84, 57, 105),
                       result = c("win", "loss", "win", "loss", "win", "loss","win", "loss","win", "loss", "win", "loss"),
                       season = c(2022, 2022, 2022, 2022, 2021, 2021, 2021, 2021, 2019, 2019, 2019, 2019))

win_loss %>% 
  group_by(season) %>% 
  ggplot(aes(x = factor(team), y = record, fill = result)) +
  geom_col(position = "dodge") + 
  geom_text(aes(label = record), vjust = -.5, position = position_dodge(.9), family = "serif") +
  labs(x = "Team", y = "Record", fill = "Game Result") +
  #ylim(0, 110) +
  theme_bw() +
  scale_fill_manual(values = c("navyblue", "cornflowerblue")) +
  theme(text=element_text(family = "serif")) +
  facet_wrap(~season)

```

### Payroll

```{r, message=FALSE, warning=FALSE, fig.align='center'}
payroll <- data.frame(team = c("TEX", "MIA", "avg", "TEX", "MIA", "avg", "TEX", "MIA", "avg"),
                      season = c(2021, 2021, 2021, 2022, 2022, 2022, 2023, 2023, 2023),
                      payroll = c(95788819, 58157900, 130886243, 150037446, 82954422, 149990188, 
                                  182832500, 81075000, 146680879))
payroll %>% 
  group_by(season) %>% 
  ggplot(aes(x = factor(team), y = payroll, fill = team)) +
  geom_col(position = "dodge") +
  labs(title = "MLB Payrolls", subtitle = "League Average, Miami Marlins & Texas Rangers", 
       x = "Team", y = "Payroll", color = "Team") +
  theme_bw() +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("slategray3", "deepskyblue", "red2")) +
  theme(text=element_text(family = "serif")) +
  theme(legend.position = "false") +
  facet_wrap(~season)
```
## Standings

### AL

**The first table is standings the day before the game of interest. The second is the standings on 9/12.**

```{r}
standings11_al <- bref_standings_on_date(date = "2022-09-11", division = "AL Overall")
standings12_al <- bref_standings_on_date(date = "2022-09-12", division = "AL Overall")

#bref_standings_on_date(date = "2022-09-11", division = "NL Overall")

# add row to show ranks
standings11_al <- standings11_al %>% 
  mutate(rank = 1:15) %>% 
  dplyr::select(rank, Tm:`pythW-L%`)
standings12_al <- standings12_al %>% 
  mutate(rank = 1:15) %>% 
  dplyr::select(rank, Tm:`pythW-L%`)

kable(standings11_al, caption = "AL - Day Before") %>% 
  kable_styling("striped") %>% 
  row_spec(12, background = "skyblue") %>% 
  row_spec(5, background = "lavender") %>% 
  row_spec(15, background = "gainsboro")
kable(standings12_al, caption = "AL - On 9/12") %>% 
  kable_styling("striped") %>% 
  row_spec(11, background = "skyblue") %>% 
  row_spec(3, background = "lavender") %>% 
  row_spec(15, background = "gainsboro")
```

Rangers went up 1 in the rankings. Toronto went up 2. Oakland stayed at 15.

### NL

**The first table is standings the day before the game of interest. The second is the standings on 9/12.**

```{r}
standings11_nl <- bref_standings_on_date(date = "2022-09-11", division = "NL Overall")
standings12_nl <- bref_standings_on_date(date = "2022-09-12", division = "NL Overall")

# add row to show ranks
standings11_nl <- standings11_nl %>% 
  mutate(rank = 1:15) %>% 
  dplyr::select(rank, Tm:`pythW-L%`)
standings12_nl <- standings12_nl %>% 
  mutate(rank = 1:15) %>% 
  dplyr::select(rank, Tm:`pythW-L%`)

kable(standings11_nl, caption = "AL - Day Before") %>% 
  kable_styling("striped") %>% 
  row_spec(12, background = "skyblue") 
kable(standings12_nl, caption = "AL - On 9/12") %>% 
  kable_styling("striped") %>% 
  row_spec(12, background = "skyblue") 
```

MIA stayed the same in the rankings

## Trevor Rogers

```{r, include=FALSE}
# rogers20 <- statcast_search(start_date = "2020-03-01", end_date= "2021-02-01", playerid = 669432)
# rogers21 <- statcast_search(start_date = "2021-03-01", end_date= "2022-02-01", playerid = 669432)
# 
# rogers22 <- statcast_search(start_date = "2022-04-01", end_date= "2022-09-30", playerid = 669432)
# rogers_dates <- fg_pitcher_game_logs(playerid = 22286, year = 2022) %>% View()
# 
# rogers_key <- data.frame(date = rogers_dates$Date)
# 
# keys <- c(663141, 661420, 662283, 661831, 663046, 661717, 662700, 662814)
# 
# mlb_pbp(663141)
```

```{r}
rogers <- data.frame(count = c("0-1", "0-2", "1-0", "1-1", "1-2", "2-0", "2-1", "2-2", "3-0", "3-1", "3-2"),
                     `K/BB` = c(9.09, 66.0, 1.24, 3.04, 14.64, 0.54, 1.27, 4.54, 0.14, 0.36, 1.44),
                     OBP = c(.280, .190, .362, .312, .207, .515, .407, .277, .762, .620, .367))

rogers %>% 
  kable(col.names = c("count", "K/BB", "OBP")) %>% 
  kable_styling("striped") 
```

66.0 is not a type, it's really that high. I can re-format this table in any way you'd like just let me know how you want it to look.

I could make this into two separate bar graphs if you'd like. They can't be on the same one since the scales are so different, and one of the K/BB values is much either than the others so it would make it hard to see all the values without using a transformation (like log). 

## Postseason

```{r}
post2022 <- mlb_schedule_postseason(season = 2022) %>%
  dplyr::select(date:game_pk, game_type, season, official_date, description, teams_away_team_id, teams_away_team_name, teams_away_score, teams_home_team_id,
                teams_home_team_name, teams_home_score)
post2021 <- mlb_schedule_postseason(season = 2021) %>%
  dplyr::select(date:game_pk, game_type, season, official_date, description, teams_away_team_id, teams_away_team_name, teams_away_score, teams_home_team_id,
                teams_home_team_name, teams_home_score)
post2020 <- mlb_schedule_postseason(season = 2020) %>%
  dplyr::select(date:game_pk, game_type, season, official_date, description, teams_away_team_id, teams_away_team_name, teams_away_score, teams_home_team_id,
                teams_home_team_name, teams_home_score)
post2019 <- mlb_schedule_postseason(season = 2019) %>%
  dplyr::select(date:game_pk, game_type, season, official_date, description, teams_away_team_id, teams_away_team_name, teams_away_score, teams_home_team_id,
                teams_home_team_name, teams_home_score)
post2018 <- mlb_schedule_postseason(season = 2018) %>%
  dplyr::select(date:game_pk, game_type, season, official_date, description, teams_away_team_id, teams_away_team_name, teams_away_score, teams_home_team_id,
                teams_home_team_name, teams_home_score)
post5 <- rbind(post2022, post2021, post2020, post2019, post2018)
```

```{r}
# rangers
# post5 %>% 
#   filter(teams_away_team_id == 140 | teams_home_team_id == 140) %>% View()
# # blue jays
# post5 %>% 
#   filter(teams_away_team_id == 141 | teams_home_team_id == 141) %>% View()
```

